---
title: "Gathering Data from ESPN's Undocumented API"
output: 
  html_document:
    includes: 
      in_header: sources.html
      after_body: footer.html
    toc: true
    toc_depth: 2
    toc_float: true
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

## Overview

I prepared a limited version of the 'Luck Rankings' for the 2018-2019 season. For this analysis I was mostly limited by the data I had available to me. I manually entered the score of each matchup into a csv and then ran analysis on those scores. This year, I wanted to do a richer level of analysis, at the player level. Therefore, I had to figure out how to programmatially get the data from the web. Luckily, there were a few reddit and blog posts that helped me out a lot (links to follow). The remainder of this post contains the code that I used to download and prepare the data. 

## Python API Downloading 

The following Python code downloads the API data into a text file which I will parse in R. 

```{python, eval = FALSE}
# imports
import requests
import json

# set values
url = "https://fantasy.espn.com/apis/v3/games/ffl/seasons/2019/segments/0/leagues/554351"

# Get cookie values from secret file
with open ("C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Cookies/espn_s2.txt", "r") as myfile:
    espn_s2=myfile.readlines()

with open ("C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Cookies/SWID.txt", "r") as myfile:
    SWID=myfile.readlines()

cookies = {"espn_s2": espn_s2[0], "SWID": SWID[0]}

# get request - league Data
r = requests.get(url, cookies=cookies)
d = r.json()
with open('C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/leagueData.txt', 'w') as outfile:
    json.dump(d, outfile)

# get request - mMatchup view
reports = ["mMatchup", "mTeam", "mMatchupScore", "kona_player_info", "mSettings"]
for report in reports:
    r = requests.get(url, params = {"view": report}, cookies=cookies)
    d = r.json()
    filename = "C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/" + report + ".txt"
    with open(filename, 'w') as outfile:
        json.dump(d, outfile)

# get request - boxscore2 (player level data - for each team)
for weeks in range(1, 11):
    r = requests.get(url + '?view=mMatchup&view=mMatchupScore', params={'scoringPeriodId': weeks, 'matchupPeriodId': weeks}, cookies=cookies)
    d = r.json()
    filename = "C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/boxscore2" + str(weeks) + ".txt"
    with open(filename, 'w') as outfile:
        json.dump(d, outfile)

# get request - kona_view (player level data for the entire year, not team specific)
for weeks2 in range(1, 11):
    r = requests.get(url + '?view=kona_player_info', params={'scoringPeriodId': weeks2}, cookies=cookies)
    d = r.json()
    filename = "C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/kona_player_info_" + str(weeks2) + ".txt"
    with open(filename, 'w') as outfile:
        json.dump(d, outfile)

```

<br><br>

## Data Munging 

<br>

### Matchup Score Data

```{r data1, eval = FALSE}
keep(thisWeek, rMarkdownWeeklyScript, sure = TRUE)

currentWeek = 10 # The latest complete week in the season. 
simulate.schedules = 1
import.player.data = 1


##########################x
## Matchup Scoring Data
##########################x
mMatchupView = fromJSON("C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/mMatchup.txt")

dat = data.frame(week       = mMatchupView$schedule$matchupPeriodId,
                 awayPoints = mMatchupView$schedule$away$totalPoints,
                 awayID     = mMatchupView$schedule$away$teamId,
                 homePoints = mMatchupView$schedule$home$totalPoints,
                 homeID     = mMatchupView$schedule$home$teamId) %>%
      group_by(week) %>% mutate(matchupID = row_number())

awayResults <- rename(dat,points = awayPoints, oppPoints = homePoints, ID = awayID, oppID = homeID)
homeResults <- rename(dat,points = homePoints, oppPoints = awayPoints, ID = homeID, oppID = awayID)
dat <- rbind(awayResults, homeResults)

```

<br>

### League Data

```{r data2, eval = FALSE}
##########################x
## League Overview Data - team names and owners
##########################x

leagueData = fromJSON("C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/leagueData.txt")

leagueDat.MASTER = data.frame(ID = leagueData$teams$id,
                       loca = leagueData$teams$location,
                       name = leagueData$teams$nickname) %>%
  mutate(fullname = paste(loca,name)) %>% select(-c(loca, name)) %>%
  mutate(fullname = ifelse(ID == 4, "Are Bonita Fish Big?", fullname)) %>% # getting rid of roland's emoji
  mutate(owner = ifelse(ID==1,  "Cameron",   ifelse(ID==3,  "KB",
                 ifelse(ID==4,  "Roland",    ifelse(ID==7,  "Carter/Ale",
                 ifelse(ID==9,  "Kolin",     ifelse(ID==10, "Luke",
                 ifelse(ID==11, "Trainor",   ifelse(ID==15, "Tex",
                 ifelse(ID==16, "Rich",      ifelse(ID==17, "KY","")))))))))))

dat.MASTER <- dat <- merge(dat, leagueDat.MASTER) %>% mutate(winner = ifelse(points > oppPoints, 1, 0))
```

<br>

### Boxscore Data

```{r data3, eval = FALSE}
##########################x
## Boxscore Data
##########################x
weekList = vector(mode = "list", length = currentWeek)

for (weeks in 1:currentWeek) {
  # read in weekly data
  filename = paste0("C:/Users/lvwilson/Documents/GitHub/lvzwilson1.github.io/Ignore/Data/boxscore2",weeks,".txt")
  boxscorejson = fromJSON(filename)
  teams <- vector(mode = "list", length = 5)
  
  for (team in 1:5) {
    offset = (weeks - 1)*5
    awayAndHomeList = list(as.list(boxscorejson$schedule$away$rosterForCurrentScoringPeriod$entries[[team+offset]]),
                           as.list(boxscorejson$schedule$home$rosterForCurrentScoringPeriod$entries[[team+offset]]))
    mMatchupViews =   list(mMatchupView$schedule$away,mMatchupView$schedule$home)
    homeandawayDats = vector(mode = "list", length = 2)
    
    for (list in 1:2) {
      teamID = mMatchupViews[[list]]$teamId[team+offset]
      nPLayers = length(awayAndHomeList[[list]]$lineupSlotId)
      TeamPlayer <- vector(mode= "list", length = nPLayers)
      
      for (playerNum in 1:nPLayers) {
        ProjIndex <- which(awayAndHomeList[[list]]$playerPoolEntry$player$stats[[playerNum]]$statSourceId == 1)
        
        if (is.null(awayAndHomeList[[list]]$playerPoolEntry$player$stats[[playerNum]]$appliedTotal[ProjIndex]) | 
            length(awayAndHomeList[[list]]$playerPoolEntry$player$stats[[playerNum]]$appliedTotal)==1) {
          proj = 0
        } else {
          proj =    awayAndHomeList[[list]]$playerPoolEntry$player$stats[[playerNum]]$appliedTotal[ProjIndex] 
        }
        
        TeamPlayer[[playerNum]] = data.frame(
          slotID = awayAndHomeList[[list]]$lineupSlotId[playerNum],
          points = awayAndHomeList[[list]]$playerPoolEntry$appliedStatTotal[playerNum],
          name = awayAndHomeList[[list]]$playerPoolEntry$player$fullName[playerNum],
          proj = proj, ID = teamID,
          possSlots = awayAndHomeList[[list]]$playerPoolEntry$player$eligibleSlots[[playerNum]]
        )  
      }
      
      homeandawayDats[[list]] <- plyr::rbind.fill(TeamPlayer)
    }
    teams[[team]] <- plyr::rbind.fill(homeandawayDats)
  }
  weekList[[weeks]] <- plyr::rbind.fill(teams) %>% mutate(week = weeks)
  print(paste(weeks))
}

# turn league wide list into league wide dataframe
playerscoresDat.MASTER = plyr::rbind.fill(weekList) 

#cleanup
rm(teams, TeamPlayer, boxscorejson, leagueData, mMatchupView, mMatchupViews, 
   weekList, homeandawayDats, awayAndHomeList, awayResults, homeResults)

```

<br><br>

## Limited Analysis

<br>

### Best Optimal and Projected Linueps

```{r analysis1, eval = FALSE}
##########################x
## Best Lineups
##########################x

'%ni%' <- Negate('%in%')
usedSlots = unique(playerscoresDat.MASTER$slotID)
teams = unique(playerscoresDat.MASTER$ID)
weeks = unique(playerscoresDat.MASTER$week)
playerscoresDatt <- playerscoresDat.MASTER %>% filter(possSlots %in% usedSlots & possSlots != 20) %>% mutate(name = as.character(name))

lineups = vector(mode = "list", length = length(teams))

for (team in teams) {
  weeklylineups = vector(mode = "list", length = length(weeks))
  
  for (weeknum in 1:currentWeek) {
    # duplicate data, need to loop over actual and proj best lineups
    temp <- playerscoresDatt %>% filter(week == weeknum & ID == team) %>% rbind(.,.) %>% mutate(scoreType = rep(c(1,2), each = n()/2)) %>%
      mutate(typeScore = ifelse(scoreType==1, points, proj))
    
    bestLists = vector(mode = "list", length = 6)
    posis1 = c(0,6,16,17); posis2 = c(4,2)
    
    # positions with one slot - QB, TE, DST, K
    for (posi in 1:length(posis1)) {
      bestLists[[posi]]   <- temp %>% filter(possSlots==posis1[posi]) %>% group_by(scoreType) %>% 
                                      arrange(desc(typeScore)) %>% slice(1)
    }
    
    # positions with one slot - WR, RB
    for (posi in 1:length(posis2)) {
      bestLists[[posi+4]] <- temp %>% filter(possSlots==posis2[posi]) %>% group_by(scoreType) %>% 
                                      arrange(desc(typeScore)) %>% slice(1:2)
    }
    
    bestplayers <- plyr::rbind.fill(bestLists)
    
    # get top flex players -  by actual scores and projected scores
    usedNamesAct  = bestplayers[bestplayers$possSlots %in% c(4,2,6) & bestplayers$scoreType == 1,3]
    topFlexAct    <- temp %>% filter(name %ni% usedNamesAct & possSlots == 23 & scoreType == 1) %>%
                              arrange(desc(typeScore)) %>% slice(1) 
    
    usedNamesProj = bestplayers[bestplayers$possSlots %in% c(4,2,6) & bestplayers$scoreType == 2,3]
    topFlexProj   <- temp %>% filter(name %ni% usedNamesProj & possSlots == 23 & scoreType == 2) %>% 
                              arrange(desc(typeScore)) %>% slice(1)
    
    # bind together dataframe of top players
    weeklylineups[[weeknum]] = rbind(bestplayers, topFlexAct, topFlexProj)
  }
  
  lineups[[which(teams == team)]] <- plyr::rbind.fill(weeklylineups)
}

lineUps.MASTER <- plyr::rbind.fill(lineups) %>% arrange(week, ID, scoreType, slotID) %>%
  group_by(week, ID, scoreType) %>% summarize(maxScore = sum(points)) %>% ungroup %>% 
  spread(scoreType, maxScore)   %>% rename(ActualMax="1", ProjScore="2") %>% inner_join(dat) %>%
  mutate(espnWinner = ifelse(ProjScore > oppPoints,1,0))

table(lineUps.MASTER$espnWinner, lineUps.MASTER$winner)

rm(bestLists, bestplayers, topFlexAct, topFlexProj, weeklylineups, lineups, temp, playerscoresDatt)

```

<br>

```{r analysis2, eval = FALSE}
############################ x
### Simulate All Schedules....
############################ x  

schedules <- vector(mode = "list", length = 10)
team_ids <- unique(dat$ID)

## Generate dataframe of all possible schedules
for(i in 1:length(team_ids)) {
  teamID <- team_ids[i]
  teams.except <- team_ids[team_ids != teamID]
  schedules[[i]] <- combinat::permn(teams.except)
  nr = length(schedules[[i]])
  
  for (j in 1:nr) {
    schedules[[i]][[j]] <- c(schedules[[i]][[j]], schedules[[i]][[j]][1:4]) 
  }
  
  schedules[[i]] <- unlist(schedules[[i]])
  schedules[[i]] <- data.frame(teamID = team_ids[i], nschedule = rep(1:362880, each = 13),
                               week = rep(1:13, times = 362880), opponentID = schedules[[i]])
  print(i)
}

all.sched.MASTER = plyr::rbind.fill(schedules)
all.sched.MASTER %<>% filter(week <= currentWeek)

rm(schedules)

## merge on points scored by each team - takes forever
scores         <- dat[,c("week", "ID", "points")] %>% rename(teamID=ID)
scoresOpponent <- dat[,c("week", "ID", "points")] %>% rename(opponentID=ID, oppPoints = points)
all.sched.MASTER <- inner_join(all.sched.MASTER, scoresOpponent) 
all.sched.MASTER <- inner_join(all.sched.MASTER, scores)
all.sched.MASTER <- mutate(all.sched.MASTER, winner = ifelse(points > oppPoints, 1, 0))
  

keep(dat.MASTER, leagueDat.MASTER, playerscoresDat.MASTER, lineUps.MASTER, 
     all.sched.MASTER, thisWeek, rMarkdownWeeklyScript, sure = TRUE)

```


